---
title: "Geography 176A Lab-02"
author: '[Johann Ayala-Vargas](https://mikejohnson51.github.io)'
subtitle: 'Lab 03: COVID-19 Pandemic'
output:
  html_document:
    theme: journal
  pdf_document: default
---
install.packages("zoo")
#installing zoo package
```{r}
library(tidyverse)
library(knitr)
library(readxl)
library(zoo)
```
#this adds all of these packages into our library for this project


#Question 1:

```{r}
url = 'https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv'

covid = read_csv(url)
```


#Importing data via url
```{r}
url = 'https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv'

covid = read_csv(url)
```



#Filtering our data to only generate results for New Cases each day to only California Counties\

#Dataframe showing Cumulative cases in descending order
```{r}
covid2 = covid %>%  
  filter(state == ("California")) %>% 
  group_by(county) %>% 
  mutate(newCases = cases - lag(cases)) %>%
  ungroup() %>% 
  filter(date == max(date))%>%
  select(state, county, cases, newCases) %>% 
  slice_max(cases, n = 5)
```

#Dataframe showing new cases in descending order
```{r}
covid3 = covid %>%  
  filter(state == ("California")) %>% 
  group_by(county) %>% 
  mutate(newCases = cases - lag(cases)) %>%
  ungroup() %>% 
  filter(date == max(date))%>%
  select(state, county, cases, newCases) %>% 
  slice_max(newCases, n = 5)
```




#shows our data table with captions for how to interpret the results. 
```{r}
knitr::kable(covid3, 
             caption = "Most New Cases California Counties",
             col.names = c("State", "County", "Cases", "New Cases"))
knitr::kable(covid2, 
             caption = "Most Cumulative Cases California Counties",
             col.names = c("State", "County", "Cases", "New Cases"))

```

#allows us to read excel files for the data we are about to import.
```{r}
library(readxl)
```


#loading the downloaded excel data
```{r}
library(readxl)
Pop <- read_excel("~/github/geog-176A-labs/data/PopulationEstimates.xls", skip = 2)
```

```{r}
Pop2 = Pop %>%
  select(fips = FIPStxt, state = State, Area_Name, pop_2019 = POP_ESTIMATE_2019) %>%
  right_join(covid, pop, by = "fips") %>%
  filter(date >= max(date)-13, state.y == "California") %>% 
  group_by(county) %>%
  mutate(cumul_per_cap = sum(cases) / pop_2019) %>%
  mutate(new_cases_per_cap = sum(cases - lag(cases), na.rm= TRUE)/ pop_2019) %>%
  select(county, cumul_per_cap, new_cases_per_cap)

(most_new_cases_per_cap = Pop2 %>%
  slice_max(new_cases_per_cap, n = 5) %>%
  select(county, cumul_per_cap))

(most_cumul_per_cap = Pop2 %>%
    slice_max(cumul_per_cap, n = 5) %>%
  select(county, cumul_per_cap))
  
library(formattable)

knitr::kable(most_new_cases_per_cap,
             caption = "Most New Cases Per Capita: California",
             col.names = c("County", "New Cases Per Capita"),
             format.args = list(big.mark = ","))

knitr::kable(most_cumul_per_cap,
             caption = "Most Cumulative Cases Per Capita: California",
             col.names = c("County", "Cumulative Cases Per Capita"),
             format.args = list(big.mark = ","))



```





```{r}
names(Pop)
dim(Pop)
nrow(Pop)
```




```{r}
#last14days =  Pop2 %>%
  #filter(date > max(date) - 13) %>%
  #group_by(county, pop_2019)%>%
  #summarise(newCases = sum(newCases)) %>%
  #ungroup() %>%
  #mutate(casePer100 = newCases / (pop_2019 / 100000)) %>%
  #filter(casePer100 <= 100) %>%
  #pull(county)

#could not get code to run
```




#Question 2

#filters the data like covid2, but now this data also includes values for New York, Florida and Louisiana

```{r}
covid3 = covid %>%  
  filter(state == c("California", "New York", "Florida", "Louisiana")) %>% 
  group_by(county) %>% 
  mutate(newCases = cases - lag(cases)) %>%
  ungroup() %>% 
  filter(date == max(date))%>%
  select(state, county, cases, newCases)
```


```{r}
  
Most_cumulative2 = covid3[order(-covid3$cases),]

Most_new2 = covid3[order(-covid3$newCases),]

covid3 %>% 
  group_by(state) %>% 
  mutate(roll_mean = rollmean(cases, 2, na.pad = T))
Most_cumulative2 = covid3[order(-covid3$cases),]


Most_new2 = covid3[order(-covid3$newCases),]


#ggplot(data = Most_new2, aes(x = newCases, y = roll_mean)) +
  #geom_point(aes(color = county, size = cases)) +
  #labs(title = "Roll Mean of Corona Virus Counts as a Fucntion of Daily New Cases Per State", x = "Daily New Cases", y = "Roll Mean", color = "", size = "Cases") +
  #facet_wrap(~state)+
  #theme_bw()

#could not get code to run

```


